---
title: "Lab 6 Comparing two means"
author: "Nathan Snell"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean difference in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

The mean difference in the number of species between areas upstream and downstream of a tributary is 1.8334. The 95% confidence interval of this mean difference is -4.587031 to 8.253697.  

### Load packages
```{r}
library(tidyverse)

fish <- read_csv("chap12q19ElectricFish.csv")
fish

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

fish_long

```


```{r}
t.test(formula= species ~ location, data= fish_long)
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

The t-test shows that there is no significant difference between the species and the location since P > 0.05 (t = 0.59249, df = 21.81, p = 0.5596). The difference between means falls in the 95% confidence interval. Therefore, we fail to reject the null hypothesis. For this reason we can say that the tributaries have no effect on the number of species of electric fish.
```{r}
t.test(formula= species ~ location, data= fish_long)
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

To complete part A and B we had to assume that the scale of assumption follows an ordinal scale, the saple of the population was collected randomly, the data follows a normal curve, and equal variance between sets exists when standard deviations of samples are about equal. As you can see in the histogram, the number of species found at heach site follows a normal curve. 
```{r}
ggplot(data = fish_long) +
  geom_histogram(mapping = aes(x = species), binwidth = 4,
                 boundary = 0, closed = "left", 
                 fill = "#C5351B", color = "black") +
  labs(x = "Species", y = "Frequency (number of species)") +
  scale_y_continuous(breaks = seq(0, 8, 2), limits = c(0, 8), 
                     expand = expansion(mult = 0)) +
  scale_x_continuous(breaks = seq(0, 36, 4), limits = c(0, 40)) +
  theme_classic() +
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1)))
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}
crabs <- read_csv("chap15q27FiddlerCrabFans.csv")
crabs
```


```{r}
crabs %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType ), 
    bins = 15, 
    alpha = 0.5,
    position= "identity",
    na.rm = TRUE)+
  labs(x= "Temperature", y="Frequency", fill="Type of Crab")+
  facet_wrap(~crabType)

crabs %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType ), 
    bins = 15, 
    alpha = 0.5,
    position= "identity",
    na.rm = TRUE)+
  labs(x= "Temperature", y="Frequency", fill="Type of Crab")
```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

The null hypothesis would be that there is no significant difference between body temperature of the crab types. The alternative hypothesis is that there is a significant difference between body temperature of the crab types.
```{r}
aov_crabs <-
  aov(bodyTemperature ~ crabType, data = crabs)
aov_crabs

summary(aov_crabs)
```

The results show that there is a significant difference of temperatures between crab types. We know this because the F value is not close to 1 and the P value is smaller than 0.05. Therefore, we can reject the null hypothesis. 
